Joe Biden Predicted to Win 59% of the Popular Vote in the 2024 U.S. Presidential Election

### 2024-03-15 paper/paper.qmd
model and paper so far

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Talia Fabregas
  - Fatimah Yunusa
  - Aamishi Sundeep Avarsekar
thanks: "Code and data are available at: https://github.com/taliafabs/US-Election-Forecast-2024.git"
date: today
date-format: long
abstract: "The 2024 U.S. Presidential Election ..."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

library(tidyverse)
library(janitor)
library(rstanarm)
library(readr)
library(ggplot2)
library(knitr)
library(arrow)
```

@@

The 2024 U.S. Presidential Election will take place on Tuesday November 5 2024. Incumbent President Joseph R. Biden Jr. will seek a second term. Former President Donald J. Trump GOP nominee.

Delegates to secure the nomination, Nikki Haley has dropped out of GOPYou can and should cross-reference sections and sub-sections. We use @citeR and

@@

@@

@@

proportion_biden_by_state
proportion_by_state

@@

```{r}

@@

The specific survey data set used is the America's Political Pulse Survey Data Week 3 2024 January 12-19. It contains the responses of the 1000 American adults interviewed between January 12-19 2024. Demographic information about respondents, including sex, age, race, home state, employment status, marital status, ... is included. Participants were als asked who they voted for in the 2016 and 2020 U.S. presidential characteristics, what their party affiliation is, and what their political ideology is.

@@

### Survey Data figures and tables go here

In order to gain a better understanding of our survey data, we conducted exploratory data analysis.

@@

@@

survey_analysis_data |>
  ggplot(aes(x = races, fill = vote24)) +
  stat_count(position = "dodge") +
  facet_wrap(facets = vars(sex)) +
  theme_minimal() +
  labs(
    x = "Race",
    y = "Number of respondents",
    fill = "Preferred Presidential Candidate"
  ) +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

```{r}
# write about America's urban-rural divide
#| echo: false
#| warning: false
#| label: fig-survey_urban_rural
#| fig-cap: The Distribution of Presidential Preferences of Respondents Living in Urban vs Rural Areas
survey_analysis_data |>
  ggplot(aes(x = urban, fill = vote24)) +
  stat_count(position = "dodge") +
  # facet_wrap(facets = vars(sex)) +
  theme_minimal() +
  labs(
    y = "Number of respondents",
    fill = "Preferred Presidential Candidate"
  ) +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

@@

```{r}
#| echo: false
#| warning: false
#| label: tbl-survey_presidential_preferences_table
#| fig-cap: Who Voters Plan to Support (Survey Data)

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-electoral-map-survey
#| fig-cap: Electoral map based solely on the survey data
library(ggplot2)
library(statebins)
survey_electoral_map <- survey_analysis_data %>% 
  group_by(state, vote_biden) %>% 
  summarise(n = n()) %>% 
  group_by(state) %>% 
  mutate(prop = n/sum(n)) %>% 
  filter(vote_biden == 1)

survey_electoral_map %>%
  mutate(statename = str_to_title(state)) %>%
  ggplot(aes(fill = prop, state = statename)) +
  geom_statebins() +
  scale_fill_gradient2(low = "#d12531", high = "#244999",
                       mid = "white", midpoint = mean(survey_electoral_map$prop)) +
  theme_classic() +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
  labs(fill = "Proportion Voting \nfor Biden",
       title = "Map of the US States that plan to support Joe Biden or Donald Trump")

@@

@fig-electoral-map-survey illustrates the proportion of voters in each U.S. state who plan to support President Biden in the 2024 Presidential Election. We can see that there is strong support for President Biden in Washington, Oregon, New York, New Jersey, Maryland, Massachusetts, and Connecticut. We also see that Oklahoma, Idaho, and Utah heavily favor former President Trump. These results are expected and there were similar margins in the 2020 election. These states have also been Dem and GOP strongholds since 1992 (or whatever year). However, this electoral map also highlights some of the limitations fo teh survey data. Wyoming and Kansas are very blue in the electoral map based solely on the survey data, however that is very far from recent election results in those states. In fact, Wyoming was the most Pro-Trump state in both the 2016 and 2020 presidential elections, with Trump receiving 2016% and 2020% of the vote in each of those elections, respectively. The survey data set includes very few examples from small states.

@@

```{r}
#| label: fig-temp1
#| fig-cap: abc
#| echo: false

@@

```{r}
#| label: fig-temp2
#| fig-cap: temp
#| echo: false
#| warning: false
#| message: false

# analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
# 
# analysis_data |> 
#   ggplot(aes(x = width, y = length)) +
#   geom_point(alpha = 0.8) +
#   theme_minimal() +
#   labs(x = "Wing width (mm)",
#        y = "Wing length (mm)")
```

@@

logistic regression

binary

predicts support for trump or biden

<!-- # Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details]. -->

```{r survey_model, echo=F, message=F, fig.cap="Results of Regression Model"}
library(rstanarm)
library(arrow)
library(dplyr)

# poststrat_analy <- arrow::read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/poststrat_analysis_data.parquet")
# 
# poststrat_data_cells <-
#   poststrat_analysis_data |>
#   count(state, sex, age_group, education_level)
# 
# poststrat_data_cells <-
#   poststrat_data_cells |>
#   mutate(prop = n / sum(n),
#          .by = stateicp)
# 
# poststrat_data_cells
```

@@

Define $y_i$ whether an individual's preferred 2024 presidential candidate is Joe Biden (1) or Donald Trump (0). The logistic regression model used to predict support for Biden using sex, age, race, highest level of education, state, and whether the individual lives in an urban or rural area is as follows:

```{=tex}
\begin{align}
Pr(vote\_biden_i =1) = logit^{-1}( \beta_0 + \beta_1 \times x_{sexi} + \beta_2 \times x_{age\_bracketi} +  \beta_3 \times x_{racesi} + \\
\beta_3 \times x_{education\_leveli} + \beta_4 \times x_{statei} + \beta_5 
\times x_{urbani} + \beta_5 \times x_{race\_blacki})
\end{align}
```
We run our model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

@@

@@

We can use maths by including latex between dollar signs, for instance $\theta$.

@@

Our results are summarized in @tbl-modelresults.

presidential_vote_model <-
  readRDS(file = here::here("models/presidential_vote_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of presidential vote based on ..."
#| warning: false

modelsummary::modelsummary(
  list(
    "Presidential Vote Model" = presidential_vote_model
  ),
  statistic = "mad",
  fmt = 2
)
```

@@

@@

>>>>

---
title: "Joe Biden Predicted to Win 59% of the Popular Vote in the 2024 U.S. Presidential Election"
subtitle: "A 2024 U.S. Election Forecast Using Multi-Level Regression with Post-Stratification (MRP)"
author: 
  - Talia Fabregas
  - Fatimah Yunusa
thanks: "Code and data are available at: https://github.com/taliafabs/US-Election-Forecast-2024.git"
date: today
date-format: long
abstract: "The 2024 U.S. Presidential Election will take place on November 5th and it will be a rematch between President Joe Biden and Former President Donald Trump. Using multi-level regression with post-stratification (MRP) we found that ... Based on our results, we can predict that Joe Biden will win 59% of the popular vote. "
format: pdf
number-sections: true
bibliography: references.bib
---

@@

library(tidyverse)
library(janitor)
library(rstanarm)
library(readr)
library(ggplot2)
library(knitr)
library(arrow)

## load data
survey_analysis_data <- arrow::read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/survey_analysis_data.parquet")

poststrat_analysis_data <- arrow::read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/poststrat_analysis_data.parquet")
```

@@

The 2024 U.S. Presidential Election will take place on Tuesday November 5 2024. Incumbent President Joe Biden will seek a second term in a rematch against his 2020 opponent and, arguably, the most polarizing figure in American politics, former President Donald Trump.

@@

Talk about polarization in a discussion point instead of introduction

@@

ESTIMAND: SUPPORT FOR BIDEN

We will use ... to estimate support for President Joe Biden in the upcoming election.

@@

@@

@@

The specific survey data set used is the America's Political Pulse Survey Data Week 3 2024 January 12-19. It contains the responses of the 1000 American adults interviewed between January 12-19 2024. Demographic information about respondents, including sex, age, race, home state, employment status, marital status, and the size and type of city that they live in is included. Participants were also asked who they voted for in the 2016 and 2020 U.S. presidential characteristics, what their party affiliation is, and what their political ideology is.

@@

Survey Data figures and tables go here

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-racetable
#| tbl-cap: Races of Survey Respondents

races <- survey_analysis_data %>% 
  group_by(races) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "survey",
         group = "races")

races <- races |>
  select(races, n, pct)

races <- races |>
  rename(Count = n) |>
  rename(Race = races) |>
  rename(Proportion = pct)

kable(races)

@@

vote2020 <- survey_analysis_data |>
  filter(presvote20post == "Joe Biden" | presvote20post == "Donald Trump")

vote2020 |>
  ggplot(aes(x = races, fill = presvote20post)) +
  stat_count(position = "dodge") +
  facet_wrap(facets = vars(sex)) +
  theme_minimal() +
  labs(
    x = "Race",
    y = "Number of respondents",
    fill = "Preferred Presidential Candidate"
  ) +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

# 73% of respondents are white
# non hispanic white makes up 57.8% of the US population

```{r}
# write about America's urban-rural divide
#| echo: false
#| warning: false
#| label: fig-survey_urban_rural
#| fig-cap: The Distribution of Presidential Preferences of Respondents Living in Urban vs Rural Areas
vote2020 |>
  ggplot(aes(x = urban, fill = vote24)) +
  stat_count(position = "dodge") +
  facet_wrap(facets = vars(sex)) +
  theme_minimal() +
  labs(
    y = "Number of respondents",
    fill = "Preferred Presidential Candidate"
  ) +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

@@

```{r}
#| echo: false
#| warning: false
#| label: tbl-popvotesurvey
#| fig-cap: Preferred 2024 Presidential Candidate (Survey Data)

@@

support_table <- tibble(
  num_biden = num_biden,
  prop_biden = num_biden / num_respondents,
  num_trump = num_trump,
  prop_trump = num_trump / num_respondents
)

kable(support_table)
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-electoral-map-survey
#| fig-cap: Electoral map based solely on the survey data
library(ggplot2)
library(statebins)
survey_electoral_map <- survey_analysis_data %>%
  group_by(state) %>%
  summarise(prop = mean(vote_biden))

survey_electoral_map %>%
  mutate(statename = str_to_title(state)) %>%
  ggplot(aes(fill = prop, state = statename)) +
  geom_statebins() +
  scale_fill_gradient2(low = "#d12531", high = "#244999",
                       mid = "white", midpoint = mean(survey_electoral_map$prop)) +
  theme_classic() +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
  labs(fill = "Proportion Voting \nfor Biden",
       title = "Map of the US States that plan to support Biden or Trump")

@@

@fig-electoral-map-survey illustrates the proportion of voters in each U.S. state who plan to support President Biden in the 2024 Presidential Election. We can see that there is strong support for President Biden in Washington, Oregon, New York, New Jersey, Maryland, Massachusetts, and Connecticut. We also see that Oklahoma, Idaho, and Utah heavily favor former President Trump. These results are expected and reflective of the 2020 election. These states have also been Dem and GOP strongholds since 1992 (or whatever year).

```{r}
#| echo: false
#| warning: false
#| message: false
#| table: tbl-electoralcollegesurvey
#| fig-cap: Electoral College based on survey data

electoral_college <- read_csv("electoral_college_votes.csv", show_col_types = FALSE)
# downloaded and cleaned it
electoral_college$state <- tolower(electoral_college$state)

electoral_college$electoral_votes <- as.numeric(electoral_college$electoral_votes)

library(dplyr)

prop_biden_by_state <- survey_analysis_data %>%
  group_by(state) %>%
  summarise(prop = mean(vote_biden))

prop_biden_by_state$state <- tolower(prop_biden_by_state$state)

merged_survey_ec <- merge(electoral_college, prop_biden_by_state, by="state")

biden_states<- merged_survey_ec |>
  filter(prop >= 0.5)

trump_states <- merged_survey_ec |>
  filter(prop < 0.5)

biden_ec_survey <- sum(biden_states$electoral_votes)
trump_ec_survey <- sum(trump_states$electoral_votes)

electoral_college_survey <- tibble(
  Biden = biden_ec_survey,
  Trump = trump_ec_survey
)

kable(electoral_college_survey)
```

However, this electoral map also highlights some of the limitations fo the survey data. Wyoming and Kansas are very blue in the electoral map based solely on the survey data, however that is very far from recent election results in those states. In fact, Wyoming was the most Pro-Trump state in both the 2016 and 2020 presidential elections, with Trump receiving 2016% and 2020% of the vote in each of those elections, respectively. The survey data set includes very few examples from small states.

There are no respondents from Vermont in the survey data, thus Vermont is missing from the electoral college map and @tbl-electoralcollegesurvey is missing the 3 electoral votes allocated to Vermont.

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-racetable2
#| tbl-cap: Races of Population Data

races2 <- poststrat_analysis_data %>% 
  group_by(races) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "survey",
         group = "races")

races2 <- races2 |>
  select(races, n, pct)

races2 <- races2 |>
  rename(Count = n) |>
  rename(Race = races) |>
  rename(Proportion = pct)

kable(races2)
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-demographics1
#| fig-cap: Voter Demographics in Survey and Post-Stratification Data Sets
gender <- survey_analysis_data %>% 
  group_by(sex) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "survey",
         group = "gender") %>% 
  rename(level = sex)

races <- survey_analysis_data %>% 
  group_by(races) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "survey",
         group = "races") %>% 
  rename(level = races)

education <- survey_analysis_data %>% 
  group_by(education_level) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "survey",
         group = "education") %>% 
  drop_na(education_level) %>% 
  rename(level = education_level)

age <- survey_analysis_data %>% 
  group_by(age_bracket) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "survey",
         group = "age") %>% 
  rename(level = age_bracket)

hispanic <- survey_analysis_data %>% 
  group_by(race_hispanic) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "survey",
         group = "hispanic") %>% 
   rename(level = race_hispanic)

gender_poststrat <- poststrat_analysis_data %>% 
  group_by(sex) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "post-stratification",
         group = "gender") %>% 
  rename(level = sex)

races_poststrat <- poststrat_analysis_data %>% 
  group_by(races) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "post-stratification",
         group = "races") %>% 
  rename(level = races)

education_poststrat <- poststrat_analysis_data %>% 
  group_by(education_level) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "post-stratification",
         group = "education") %>% 
  drop_na(education_level) %>% 
  rename(level = education_level)

age_poststrat <- poststrat_analysis_data %>% 
  group_by(age_bracket) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "post-stratification",
         group = "age") %>% 
  rename(level = age_bracket)

state_poststrat <- poststrat_analysis_data %>% 
  group_by(state) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "post-stratification",
         group = "state") %>% 
   rename(level = state)

hispanic_poststrat <- poststrat_analysis_data %>% 
  group_by(race_hispanic) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "post-stratification",
         group = "hispanic") %>% 
   rename(level = race_hispanic)

## combine the demographic variables into a single data set

demographic_variables <- rbind(gender, gender_poststrat, 
                               age, age_poststrat, 
                               races, races_poststrat,
                               hispanic, hispanic_poststrat,
                               education, education_poststrat)

demographic_variables |> ggplot(aes(as.factor(level), pct, group=as.factor(type), linetype = as.factor(type))) +
  geom_line() + facet_grid(~group, scales = "free") +
  theme(axis.text.x = element_text(angle=70, size = 6, hjust = 1), legend.position = "bottom") +
  labs(x = "Category", y = "Proportion", linetype = "data set",
       title = "Voter Demographics") +
  scale_y_continuous(labels = scales::percent) +
  geom_hline(yintercept = 0.5, alpha = 0.35)

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-stateprops
#| tbl-cap: Survey and Post-Stratification Data Proportion of Voters by State

survey_analysis_data$state <- tolower(survey_analysis_data$state)
state <- survey_analysis_data %>%
  group_by(state) %>%
  summarise(n = n()) %>%
  mutate(pct = n/sum(n), type = "survey",
         group = "state") %>%
   rename(level = state)

state_poststrat <- poststrat_analysis_data %>%
  group_by(state) %>%
  summarise(n = n()) %>%
  mutate(pct = n/sum(n), type = "post-stratification",
         group = "state") %>%
   rename(level = state)

voters_by_state <- rbind(state, state_poststrat)

voters_by_state |> ggplot(aes(as.factor(level), pct, group=as.factor(type), linetype = as.factor(type))) +
  geom_line() + facet_grid(~group, scales = "free") +
  theme(axis.text.x = element_text(angle=70, size = 6, hjust = 1)) +
  labs(x = "State", y = "Proportion", linetype = "Data Set")+
       # title = "Proportion of Voters by State") +
  scale_y_continuous(labels = scales::percent)
```

@@

We will use multilevel regression with post-stratification (MRP) for our analysis. MRP is ...

Steps for MRP are:

@@

To predict whether a person will voter for Joe Biden or Donald Trump, we built a logistic regression model using data from the America's Political Pulse Survey [@americaspoliticalpulse] and post-stratified it using ACS census data [@citeIPUMS]. Logistic regression only works for binary classification, therefore we created a vote_biden variable during our data cleaning process which is equal to 1 if the person will vote for Joe Biden and 0 if the person will vote for Donald Trump.

The logistic regression model used to predict support for Biden using sex, age, race, highest level of education, state, and whether the individual lives in an urban or rural area is as follows:

<small>

$$
y_i = \beta_0 + \beta_1 \times \text{sex}_i + \beta_2 \times \text{age\_bracket}_i + \beta_3 \times \text{race}_i + \beta_4 \times \text{education\_level}_i + \beta_5 \times \text{state}_i + \beta_6 \times \text{urban}_i 
$$

</small>

logistic activation <smalL> $$
Pr(vote\_biden=1) = logit^{-1}(\beta_0 + \beta_1 \times \text{sex}_i + \beta_2 \times \text{age\_bracket}_i + \beta_3 \times \text{race}_i + \beta_4 \times \text{education\_level}_i + \beta_5 \times \text{state}_i + \beta_6 \times \text{urban}_i )
$$ </small>

another try <small>

$$
Pr(vote\_biden=1) = \text{logit}^{-1}(\beta_0 + \beta_1 \text{sex}_i + \beta_2  \text{age\_bracket}_i + \beta_3 \text{race}_i + \beta_4 \text{education\_level}_i + \beta_5 \text{state}_i + \beta_6  \text{urban}_i )
$$

</small>

We run our model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

@@

Concerns about model fit

@@

# overall support
# support by state
# electoral college map

# load the model with readRDS
presidential_vote_model <-
  readRDS(file = here::here("models/presidential_vote_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelsummary
#| tbl-cap: "Coefficient estimates"
#| warning: false

# looking at coefficient estimates using modelsummary()
modelsummary::modelsummary(presidential_vote_model)

@@

```{r}
#| eval: true
#| echo: false
#| warning: false
#|abel: tbl-supporttbl
#| tbl-cap: "Estimates for support for Biden in each state"

library(arrow)
library(broom.mixed)
library(gutenbergr)
library(haven)
library(knitr)
library(labelled)
library(modelsummary)
library(rstanarm)
library(tidybayes)
library(tidyverse)

# get proportions by demographic, then find proportion of each cell relative to population
props <- poststrat_analysis_data %>%
  group_by(state, races, age_bracket, sex, education_level, urban) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  group_by(state) %>%
  mutate(prop = n / sum(n))

poststrat_data_cells <-
  poststrat_analysis_data |>
  count(state, sex, age_bracket, races, education_level, urban)

poststrat_data_cells <-
  poststrat_data_cells |>
  mutate(prop = n / sum(n),
         .by = state)

poststrat_data_cells <- poststrat_data_cells %>%
  mutate(state = tolower(state))

biden_support_by_state <-
  presidential_vote_model |>
  add_epred_draws(newdata = poststrat_data_cells) |>
  rename(support_biden_predict = .epred) |>
  mutate(support_biden_predict_prop = support_biden_predict * prop) |>
  ungroup() |>
  summarise(support_biden_predict = sum(support_biden_predict_prop),
            .by = c(state, .draw)) |>
  summarise(
    mean = mean(support_biden_predict),
    lower = quantile(support_biden_predict, 0.025),
    upper = quantile(support_biden_predict, 0.975),
    .by = state
  )

kable(biden_support_by_state)
```

electoral college

```{r}
# googled the electoral college and made a csv
electoral_college <- read_csv("electoral_college_votes.csv", show_col_types = FALSE)
# downloaded and cleaned it

electoral_college$electoral_votes <- as.numeric(electoral_college$electoral_votes)

merged_ec <- merge(biden_support_by_state, electoral_college, by = "state")

biden_states_mean <- merged_ec |>
  filter(mean >= 0.5)

trump_states_mean <- merged_ec |>
  filter(mean < 0.5)

biden_ec_mean <- sum(biden_states_mean$electoral_votes)
trump_ec_mean <- sum(trump_states_mean$electoral_votes)

biden_states_mean <- merged_ec |>
  filter(lower >= 0.5)

trump_states_mean <- merged_ec |>
  filter(lower < 0.5)

biden_ec_lower<- sum(biden_states_mean$electoral_votes)
trump_ec_lower <- sum(trump_states_mean$electoral_votes)

The estimates for the support for Biden in each state are shown in @tbl-supporttbl

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| #| label: tbl-popvotepost
#| tbl-cap: "Estimated Popular Vote"
biden_support_overall <- presidential_vote_model %>%
  add_epred_draws(newdata = poststrat_data_cells) %>%
  rename(support_biden_predict = .epred) %>%
  mutate(support_biden_predict_prop = support_biden_predict * prop) %>%
  ungroup() %>%
  summarise(
    "Lower Estimate" = quantile(support_biden_predict, 0.05),
    "Mean Estimate" = mean(support_biden_predict),
    "Upper Estimate" = quantile(support_biden_predict, 0.95)
  )

kable(biden_support_overall)
```

```{r}
#| eval: true
#| echo: false
#| warning: false
#|abel: tbl-supporttbl
#| tbl-cap: "Estimates for support for Biden in each state"

library(arrow)
library(broom.mixed)
library(gutenbergr)
library(haven)
library(knitr)
library(labelled)
library(modelsummary)
library(rstanarm)
library(tidybayes)
library(tidyverse)

# get proportions by demographic, then find proportion of each cell relative to population
props <- poststrat_analysis_data %>%
  group_by(state, races, age_bracket, sex, education_level, urban) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  group_by(state) %>%
  mutate(prop = n / sum(n))

poststrat_data_cells <-
  poststrat_analysis_data |>
  count(state, sex, age_bracket, races, education_level, urban)

poststrat_data_cells <-
  poststrat_data_cells |>
  mutate(prop = n / sum(n),
         .by = state)

poststrat_data_cells <- poststrat_data_cells %>%
  mutate(state = tolower(state))

biden_support_by_state <-
  presidential_vote_model |>
  add_epred_draws(newdata = poststrat_data_cells) |>
  rename(support_biden_predict = .epred) |>
  mutate(support_biden_predict_prop = support_biden_predict * prop) |>
  ungroup() |>
  summarise(support_biden_predict = sum(support_biden_predict_prop),
            .by = c(state, .draw)) |>
  summarise(
    mean = mean(support_biden_predict),
    lower = quantile(support_biden_predict, 0.025),
    upper = quantile(support_biden_predict, 0.975),
    .by = state
  )

kable(biden_support_by_state)
```

visualization here

```{r}
#| echo: false
#| warning: false
#| eval: true
#| label: fig-estimatedpropsupport
#| fig-cap: "Estimated proportion of each state voting for Biden in 2024 (Post-Stratification)"

library(scales)  
biden_support_by_state |>
  ggplot(aes(y = mean, x = fct_reorder(state, mean))) +
  geom_point(color = ifelse(biden_support_by_state$mean > 0.5, "blue", "red")) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0, color = ifelse(biden_support_by_state$mean > 0.5, "blue", "red")) +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  labs(
    x = "State",
    y = "Estimated proportion support for Biden"
  ) +
  theme_classic() +
  scale_y_continuous(labels = scales::percent)+  # Setting y-axis to display percentages
  coord_flip() +
  theme(axis.text.y = element_text(size = 6)) 
```

compared to survey data shown in @fig-estimatedpropsupport2

```{r}
#| echo: false
#| eval: true
#| warning: false
#| label: fig-estimatedpropsupport2
#| fig-cap: "Estimated proportion of each state voting for Biden in 2024 Post-Stratification vs Survey Data"

survey_analysis_data$state <- tolower(survey_analysis_data$state)
biden_support_by_state |>
  ggplot(aes(y = mean, x = fct_reorder(state, mean))) +
  geom_point(color = ifelse(biden_support_by_state$mean > 0.5, "blue", "red")) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0, color = ifelse(biden_support_by_state$mean > 0.5, "blue", "red")) +
  geom_point(
    data = survey_analysis_data |>
      summarise(n = n(),
                .by = c(state, vote_biden)) |>
      mutate(prop = n / sum(n),
             .by = state) |>
      filter(vote_biden == 1),
    aes(y = prop, x = state),
    color="#A9A9A9"
    ) +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  labs(
    x = "State",
    y = "Estimated proportion support for Biden"
  ) +
  theme_classic() +
  scale_y_continuous(labels = scales::percent)+ # Setting y-axis to display percentages
  coord_flip() +
  theme(axis.text.y = element_text(size = 5)) 
```

electoral map

```{r}
#| echo: false
#| eval: true
#| warning: false
#| label: fig-electoralmappoststrat
#| fig-cap: "Electoral map based on post-stratification data"
library(ggplot2)
library(statebins)
survey_electoral_map <- biden_support_by_state %>%
  mutate(prop = mean)

survey_electoral_map %>%
  mutate(statename = str_to_title(state)) %>%
  ggplot(aes(fill = prop, state = statename)) +
  geom_statebins() +
  scale_fill_gradient2(low = "#d12531", high = "#244999",
                       mid = "white", midpoint = mean(survey_electoral_map$prop)) +
  theme_classic() +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
  labs(fill = "Proportion Voting \nfor Biden",
       title = "Map of the US States that plan to support Biden or Trump (Post-Stratification Data)")
```

type here

@@

## Data Cleaning

@@

type something here

```{r}
#| eval: true
#| echo: false
#| warning: false
#| label: fig-coefficientestimatesplot
#| layout-ncol: 2
#| fig-cap: "Coefficient Estimate"

# coefficients|> ggplot(aes(estimate, term)) +
#   geom_point() +
#   geom_errorbar(aes(xmin = conf.low, xmax = conf.high)) +
#   labs(title = "Coefficient Estimates",
#        x = "Estimate", y = "Coefficient") +
#   theme(axis.text.y = element_text(size = 5))
```
